---
title: "Linear models"
output: github_document
date: "2024-10-30"
---

## Getting Started! 

In general, remember linear regression (lm) is for continuous variables and logistic regression (generalized linear model (glm)) is for categorial variables. 

So I put a couple of packages into my library so I can do my work and the functions I want will work. The knitr used to specify information about graphics. I also used the theme_set to tell the code to make plots black and white and where to put the legend, respectively. Viridis is a color scheme and I told ggplot to use that palette. 

```{r setup}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = 0.6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Import Data

So we open this data set about NYC AirBnBs. You can always view the file first using `view(nyc_airbnb)`. We changed the rating from a scale of 10 to 1-5 by dividing each number in half. We renamed a column borough to neighborhood gorup. We removed Staten Island from borough column, and then we only output price, stars, borough, neighborhood and room type from the listed columns( the original doc had info about latitude and longitude etc.-- from here: https://p8105.com/dataset_airbnb.html) 

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)

nyc_airbnb
```

## Fit a model 

Good! so now that we have our data set and the variables that matter, we can create a model for the variables of interest. Here I will create a model that considers how price relates to the rating and the borough where the airbnb is located. 

```{r}
nyc_airbnb %>% 
  ggplot(aes( x= stars, y= price, color = borough)) +
  geom_point()
```

Ok so no we know what the data looks like. We tell R to get the linear model equation, since the ratings, price etc are all continuous variables. We can set that function equal to fit.Note! The borough options are Bronx, Brooklyn, Manhattan, Queens, and I think it might be because Bronx is entered first, OR it could be alphabetical order, but the regression is comparing all other boroughs to Bronx.  

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
fit
```

Then we can also run the summary function to get more info like degrees of freedom. The F statistic also tells us about the overall goodness of fit. 

```{r}
summary(fit)
```
I can also pull out only the coefficient from the summary function 
```{r}
summary(fit)$coef
```
You can also use the broom package to get a more condensed description. I can see the p value for each variable. We know that the regression is using Bronx as a standard and comparing the other values to this. Thus, what the coefficient estimate means is that staying in brooklyn is 40 dollars more expensive relative to the bronx, manhattan is 90 etc. ALSO notice this tells you that queens is not significantly influencing the price-- its p value is 0.145!

```{r}
broom::tidy(fit)
```
Now usually, people are only interested in the estimate and p value--- there is a way to put that out directly. I'm telling it to only select the 2 columns I care about; then I'm telling it that when it looks in the term column above and sees something that starts (^) with borough, capitalize it, add a colon and add some space before the next character appears. 

```{r}
broom::tidy(fit) %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: "))
```
Then if we wanted to see this in an actual table, we can use the `kable()` function that's used in r markdown files. Remember the knitr function is about displaying data. I can display the entire estimate and P values, or I can say hey I'm only interested in the first 3 decimals for both the estimate and p.value.

```{r}
broom::tidy(fit) %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
knitr::kable(digits =3)
```

Remember the `factor` function? Factor takes a string and attaches a numeric value to it based on its position. 

## How to change the reference category for your covariates!

Recall, I mentioned that the Bronx was automatically set as the reference category. And that's because the linear model by default, treats the variables not as characters but factors, meaning it takes a character and attaches a numeric value to it based on its position. So it uses whatever is first as the reference. 

Well! I can decide which variable should be the reference. 


## First, Let's say I want the reference column to be the most frequent borough

Here I take the nyc_airbnb modified data set from above and tell it to organize the borough column by which borough is more frequent, and also to reorder the room type by which is more frequent. So that the most frequent borough appears first. And that's what we see in the plot below.

```{r}
nyc_airbnb = 
nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))
```

Now let's see what it did to the data using the plot
```{r}
nyc_airbnb %>% 
  ggplot(aes( x= stars, y= price, color = borough)) +
  geom_point()
```
Now I rereun the regression.

I've already defined how borough should be analyzed. So note code earlier in this segment must be run prior to this for it to work. See now I have Bronx listed as a variable, and as expected, Bronx will be 90 dollars cheaper relative to Manhattan (the opposite of what we saw above). 

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

broom::tidy(fit)
```
## Diagnostics: starting with looking at residuals. 

I use the `add_residuals` function in the modelr package to add a column of the residuals to my condensed data frame, fit. 
The residual tells you the difference between your value of interest (let's say price) and where it is relative to what your model would predict. So it shows that the 99$ private room in the Bronx is about 9 dollars more expensive than my model predicts. 

It's good to look at residuals because that gives us an idea as to how well the model fits the data and if there are outliers. The taller the line/size of the squiggly space (maybe?), the more outliers there are. I reset the limits of the y axis for better visualization. 

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) + geom_violin() +
  ylim(-500,1500)


```

I can also plot the residuals for each borough and have a better idea of where the model might be failing some values--- you want to see where majority of values are-- that means they are well represented by the model. For values that are somewhat outliers, like there are a lot in manhattan, see the extra data points further away.I'm telling it to plot each borough separately (using the facet function-- I think it means all rows) and plot the residuals for each star rating 

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) + geom_point() +
  facet_wrap(. ~borough)
```

##Comparing models

So it's important to be sure that your complex model is complex because it represents the data well and is better than a simpler model (with fewer indepdent variables). The standard for comparing models is using an anova test. The original linear model (`fit = lm(price ~ stars + borough, data = nyc_airbnb)` ) looked at stars and borough as the independent variables. But let's say my null hypothesis is borough doesn't have an influence. The P value is get from the anova will tell you if there's a significant difference in the more complex plot you have, and if so, that means accounting for those extra variables DOES make a difference in how well the model estimates the effects. 
It's a little hard to understand/read at first, so I can add the `broom:: tidy()` function to improve readability. 

```{r}
fit_null = lm(price ~ stars, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough, data = nyc_airbnb)
anova(fit_null, fit_alt) %>% 
  broom:: tidy()
```

## How to create models that further stratify an indepedent variable of interest using nested data 

So this is important because let's say I want to know not only whether star rating influences price, but what if it's star rating within in each borough. I can tell the code to consider the interaction of stars and borough (and in stats means multiply). Importantly you can see p values for each interaction. 

```{r}
fit = lm(price ~ stars* borough + room_type * borough, data = nyc_airbnb) 

broom:: tidy(fit)
```

## Nesting data 

So I'm telling it to take this data and form a list of data frames. Truly not quite sure what this code is doing in terms of the -borough bc that means exclude borough. 

```{r}
nyc_airbnb %>% 
  nest(data = -borough)

```

Now I'm going to create a model for each borough. Again, not quite sure what syntax needs are for the map function. Data refers to the data column.

```{r}
nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars, data = .x))
  ) 
```

I can pull out the models column to see the coefficients. The map function takes an input (.x = the information of interest, which is data, then applies a function to elements in the list, which in this case is the linear model, and returns an object of the same length as the input, which is what I think data =.x does).

```{r}
nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars, data = .x))
  ) %>% 
  pull(models)
```

I can apply the tiny data and add a results column of the broom tidy function that provides the stats we care about. 

```{r}
stats_of_interest = nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars, data = .x)),
    results = map(models, broom::tidy)
  )
stats_of_interest
```

Now I can unnest it to show me columns and rows. I removed the stars columns

```{r}
stats_of_interest = nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars, data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
   select(-data, -models) %>% 
  unnest(results)
stats_of_interest
```

I can remove the stars column. 

```{r}
stats_of_interest = nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars, data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
   select(-data, -models) %>% 
  unnest(results) %>% 
 filter(term == "stars")
stats_of_interest
```


